A Critical Review of Sophistication and Algorithmic Model Selection

### 2015-01-06 Facticity/Facticity/paper.tex
...

\section{Preliminaries, definitions and desiderata}

@@

\section{Measures of meaningful information}

\subsection{Koppel's Sophistication}

\begin{itemize}
\item Model class is $\T$
\item Threshold function $\tau(x)=L^\M(x)+c$
\item Description method for models is required to be prefix-free, but
  can otherwise be chosen arbitrarily.
\item Koppel uses monotone complexity, and monotone functions and discusses 
  infinite strings, TODO figure out if and how this affects our results.
  \item Koppel has a `machine-independence result', but not one that does what we want.
\end{itemize}

We start with sophistication as originally introduced by Koppel and Atlan \cite{koppelSoph1988,koppel1991almost}. Their sophistication is built on a two-argument universal Turing Machine $U(i, y) = \phi_p(y)$. They consider a pair $(i, y)$ a representation of a string $x$ if (a) $\phi_i$ is total, and (b) $\phi_i(y)$ has $x$ as a prefix.\footnotemark

This construction shows one of the first pitfalls of sophistication: using an inefficient representation for the model. We show that unless 
\begin{lemma}
Koppel and Atlan's sophistication is not both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{lemma}
\begin{proof}
Assume that the sophistication is $\gamma$-invariant, for some $\gamma(n) \in o(n)$. We will show that this implies that it is bounded.

Let $\num{\phi}$ be the canonical acceptable numbering. Choose $\alpha < 1$ so that $\gamma(i) < \alpha i$ for sufficiently large $i$. Let $s: \N \to \B$ be a computable function such that $|s(i)| = \left\lceil(1-\alpha)^{-i}\right\rceil$.

We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
\[
\psi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s(2i) \\
	\infty\;\text{otherwise}
\end{cases}
\]
and 
\[
\xi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s(2i+1) \\
	\infty \;\text{otherwise}\p
\end{cases}
\]
Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations. Choose any $x$ and assume w.l.o.g. that $\s_\psi(x) < \s_\xi(x)$. By construction of the numberings $\psi$ and $\xi$, we have $\s_\psi(x)/ (1-\alpha)  \leq \s_\xi(x) + 1$, which gives us:
\begin{align*}
	\s_\xi(x) - \s_\psi(x) &\geq \alpha\, \s_\xi(x) - 1\\
	& \in \Omega(\s_\xi(x)) - 1\\
\end{align*}
while, by our original assumption $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, which contradicts the bound above.
\end{proof}

The main trick we use in this proof is that we can inflate the indexes arbitrarily. In short, the representation of the models is highly inefficient. For the definition of Kolmogorov complexity, this is not a problem, since we can switch to another universal machine at only a constant cost, but for the definition of sophistication, we must make sure that both parts of our two-part code are efficient.

We define the complexity of a given model as its shortest effective description: a description that can be used to compute the function for a given input:

\begin{definition}
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C(f) &:= \min\{C(i):\phi_i=f\}\\
    K(f) &:= \min\{K(i):\phi_i=f\}.
  \end{align*} \label{definition:model-complexity}
\end{definition}
These definitions also appear in \cite{} and \cite{}, but are not discussed in depth. We will show that these properties are still invariant up to a constant (ie. complexity is a property of a function):
\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings. There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
\begin{align*}
K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
&= K^\psi(f).
\end{align*}
We can reverse $\phi$ and $\psi$ without loss of generality to achieve the same inequality the other way around, completing the proof.
\end{proof}

This lemma shows that the complexity of a function depends on the choice of
enumeration by no more than a constant term. Note that this implies
that for any enumeration $\psi_1,\psi_2,\ldots$, the complexity $K(\psi_i)$
is less than the literal description length of the index
$|\bar\imath|$ up to a constant. For some numberings, there is a gap
between these two description methods; in those cases, the index is an
inefficient representation of the corresponding function. However it
is possible to construct numberings that are \emph{faithful} in the
sense that no such gap exists:

\begin{definition}[Faithful Numbering]\label{def:faithful}
  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
  functions is \emph{faithful} if there is a constant $c$ such that
  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$.
\end{definition}

\begin{lemma}
  There is a faithful acceptable numbering.
\end{lemma}
\begin{proof}
Let $i_\tn{div}(y)$ be an index such that $\phi_{i_\tn{div}}(y)=\infty$ for all $y$. Define
  \[\psi_q=\begin{cases}
    \phi_{\phi_i(p)}&\tn{if $q$ can be written as $\bar\imath p$ and $\phi_i(p)<\infty$,}\\
    \phi_{i_\tn{div}}&\tn{otherwise.}\end{cases}
  \]
  To show that $\psi$ is faithful, pick any function $f$. Then
\[\begin{split}
C(f)&=\min\{C(i):\phi_i=f\}\\
&=\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\}\\
&=\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}\\
&=\min\{|\bar a b|:\psi_{\bar a b}=f\}.
\end{split}\]
This shows there is a sufficiently small $\psi$ index.

To show that $\psi$ is acceptable, let $\phi_j$ denote the identity
function. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
using the computable function $r(i)=\bar\jmath i$, so that
$\psi_{r(i)}(y)=\psi_{\bar\jmath i}(y)=\phi_i(y)$. For the reverse,
define $\phi_v(\bar\imath p, y)=\phi_{\phi_i(p)}(y)$. For fixed
$\bar\imath p$, the 
$s^n_m$-theorem (see~\cite{TODO}) states that we can compute the $h$
such that $\phi_h(y)=\phi_v(\bar\imath p,y)$. Let $h(\bar\imath p)$
denote this index as a function of the program; further define
$h(q)=i_\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
construction $h$ is total and computable. To check that the mapping
returns the correct function, rewrite $\phi_{h(\bar\imath
  p)}(y)=\phi_v(\bar\imath p,y)=\phi_{\phi_i(p)}(y)=\psi_{\bar\imath p}(y)$.
\end{proof}

\begin{lemma}
  If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
\end{lemma}
\begin{proof}
Let $\bar\imath p$ be any two-part representation for the data $x$, i.e. $\phi_i(p)=x$. Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$. We compare the lengths of these two representations. Note that $|i^*|=K^\psi(i)$. Therefore,
\[
|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
\]
By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$. From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity. Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$. 
\end{proof}
Note that the condition for this lemma applies to all known partial recursive prefix functions:

@@

Subsequent publications suggest that all ensemebke

\begin{theorem}[Reduction of Effective Complexity to Structure
  Function]
  There is a constant $c$ such that, for every representation of $x$
  consisting of an ensemble $E$ with $x\in E$ with entropy $H$,
  yielding a total information of $K(E)+H$, there is a finite set $S$
  with $x\in S$ such that $K(S)\le K(E)+c$ and $\log|S|\le H$.
\end{theorem}
\begin{proof}
  First note that although the ensemble $E$ may itself be infinite, it
  is ``typical'', and therefore a valid candidate representation, for
  only a finite number of its elements, namely the set $S=\{y\in E:-\log
  P(y)\le H\} = \{y:P(y)\ge 2^{-H}\}$. Since the sum of probabilities must
  converge, this set cannot be infinite: since $1\ge\sum_{y\in S} P(y)\ge
  \sum_{y\in S}2^{-H} = |S|2^{-H}$ so $|S|\le 2^H$.

\end{proof}

@@

@@

\begin{theorem}
There are infinitely many $x$ with 
\begin{enumerate}
  \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
  \item $S^\T_{\tau^\circ}(x) \geq K(x)/6$. \label{eq:hondevoer}
\end{enumerate}
\end{theorem}
\begin{proof}
We continue from the proof of Theorem~\ref{theorem:structure-function-is-not-bounded}. Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$. Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{theorem:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$. Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
\end{proof}
Note that we favoured simplicity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.
\section{A world without sophistication}

@@

\pb{Procedural view: we cannot between dominating models. This may not be the aim of algorithmic statistics, but it suggests that it's kind of unreasonable to ask for model selection without repeated samples, or assumptions on the model class.} 
\pb{The problem with making a fixed choice.}
\pb{The existence of balanced strings.}
\pb{Total functions. They ruin efficiency and push sophistication into a land of unprovable claims.}

@@

\section{Stuff to move up or delete}

>>>>

\subsection{Preliminaries, definitions and desiderata}

@@

\section{Inefficient indices}

We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and and input. Can we assume that such a description, that is as efficient as the Kolmogorov complexity, creates a good separation into the structural part of the data and the noise?

The first problem that arises, is that allowing any acceptable numbering can lead to inefficient model representations.

\pb{Switch naar two-part perspective}
\begin{lemma}
Let $\phi$ be some acceptable numbering, and $\s_\phi(x)$ be any function which gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$.

No such function can be $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{lemma}
\begin{proof}
Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.

Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.

We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
\[
\psi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s_{2i}\\
	\infty\;\text{otherwise}
\end{cases}
\]
and 
\[
\xi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
	\infty \;\text{otherwise}\p
\end{cases}
\]
Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations. Choose any $x$ and assume w.l.o.g. that $\s_\psi(x) < \s_\xi(x)$. By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
\begin{align*}
	\s_\xi(x) - \s_\psi(x) &\geq \frac{1}{2}\, \s_\xi(x)  \in \Omega(\s_\xi(x))\p 
\end{align*}
Since, by assumption, $\s_\xi$ can be arbitrarily large and $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, we have a contradiction.
\end{proof}

In this proof, we deliberately use ridiculously inefficient index functions. Obviously one could simply pick a more efficient index; the point is that many definitions of sophistication to not mention any such restriction. As we shall see in the next, any inefficiency in the index will lead to a lack of invariance in sophistication.

The treatments that fall prey to this problem are Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost},  and several of its variants (\cite{antunes2009sophistication,antunes2013sophistication}).

Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a UTM with an efficient index at only a constant penalty (by the Invariance Theorem \cite{li1993introduction}). But this robustness property does \emph{not} carry over to definitions of sophistication, if we are interested in separating the model and its input, we must represent the model efficiently.

Another perspective ion this problem is that the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content. In order to obtain a more robust measure of model complexity, we describe a natural extension of Kolmogorov complexity to functions, and show that this retains invariance.

\begin{definition}
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\\
    K_\phi(f) &:= \min\{K(i):\phi_i=f\}.
  \end{align*} \label{definition:model-complexity}
\end{definition}
These definitions also appear in \cite{grunwald2004shannon} and \cite{vitanyi2004meaningful}. We will show that these values are invariant up to a constant (ie. complexity is an objective property of a function):
\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings. There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
\begin{align*}
K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
&= K^\psi(f).
\end{align*}
We can reverse $\phi$ and $\psi$ to achieve the same inequality the other way around, completing the proof.
\end{proof}
This lemma shows that the complexity of a function depends on the choice of
enumeration by no more than a constant term.

There are two ways to use this notion of model complexity for more robust attempts to define sophistication, that differ in the way the two-part code is conceptualized. Both approaches are used in the literature, which is therefore sometimes confusing.
\begin{enumerate}
  \item On the one hand one can stick with Koppel's approach, but repair it by imposing the additional requirement that the used length of the indices used is equal to $C(\phi_i)$, and that the numbering is still acceptable. We construct such an index below. This approach is taken by Adriaans \cite{adriaans2012facticity}. In this case, the total code length is simply $C(x)$, and the sophistication is the length of the model index within this code. This approach is the most direct formalization of the intuition that the two part code used within the definition of Kolmogorov complexity separates structure from noise.
  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity as a black box to measure the complexity of the model, which is then the first part of a two-part code describing the data. With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$. The difference is that in this approach, we can construct everything on the basis of a single, unconstrained, acceptable numbering, without using additional contraints. It is also more natural in this setting to limit the model class of allowed functions. This approach is used for Kolmogorov's structure function \cite{cover1985kolmogorov}, Algorithmic Statistics \cite{gacs2001algorithmic}, Vi\'anyi's sophistication \cite{vitanyi2004meaningful} and Effective complexity \cite{gellmann1996information}.
\end{enumerate}

To show that the two approaches are equivalent with the correct restriction on the index, we must show that there exists a numbering for which the length of the index of a model is equal to its complexity, and that such a numbering is acceptable.

\begin{definition}[Faithful Numbering]\label{def:faithful}
  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
  functions is \emph{faithful} if there is a constant $c$ such that
  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$.
\end{definition}
This definition is a refinement of the one used in \cite{adriaans2012facticity}. That publication defines an \emph{index function}, rather than a numbering, and notes that it is not computable. Since we use numberings, we only need to show that they are acceptable, ie. we only need to be able to unpakc the efficient description, so the incomputability of the index function is no issue:
\begin{lemma}
  There is a faithful acceptable numbering.
\end{lemma}
\begin{proof}
Let $i_\tn{div}(y)$ be an index such that $\phi_{i_\tn{div}}(y)=\infty$ for all $y$. Define
  \[\psi_q=\begin{cases}
    \phi_{\phi_i(p)}&\tn{if $q$ can be written as $\bar\imath p$ and $\phi_i(p)<\infty$,}\\
    \phi_{i_\tn{div}}&\tn{otherwise.}\end{cases}
  \]
  To show that $\psi$ is faithful, pick any function $f$. Then
\[\begin{split}
C(f)&=\min\{C(i):\phi_i=f\}\\
&=\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\}\\
&=\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}\\
&=\min\{|\bar a b|:\psi_{\bar a b}=f\}.
\end{split}\]
This shows there is a sufficiently small $\psi$ index.

To show that $\psi$ is acceptable, let $\phi_j$ denote the identity
function. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
using the computable function $r(i)=\bar\jmath i$, so that
$\psi_{r(i)}(y)=\psi_{\bar\jmath i}(y)=\phi_i(y)$. For the reverse,
define $\phi_v(\bar\imath p, y)=\phi_{\phi_i(p)}(y)$. For fixed
$\bar\imath p$, the 
$s^n_m$-theorem \cite{kleene193notation} states that we can compute the $h$
such that $\phi_h(y)=\phi_v(\bar\imath p,y)$. Let $h(\bar\imath p)$
denote this index as a function of the program; further define
$h(q)=i_\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
construction $h$ is total and computable. To check that the mapping
returns the correct function, rewrite $\phi_{h(\bar\imath
  p)}(y)=\phi_v(\bar\imath p,y)=\phi_{\phi_i(p)}(y)=\psi_{\bar\imath p}(y)$.
\end{proof}

Unfortunately, even with a faithful index, we can fail, if the prefix-function used to separate the model from its input is inefficient:

\begin{lemma}
Let $L(\phi) = |\bar\imath|$ If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
\end{lemma}
\begin{proof}
Let $\bar\imath p$ be any two-part representation for the data $x$, i.e. $\phi_i(p)=x$. Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$. We compare the lengths of these two representations. Note that $|i^*|=K^\psi(i)$. Therefore,
\[
|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
\]
By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$. From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity. Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$. 
\end{proof}

Note that the condition for this lemma applies to all known partial recursive prefix functions:

@@

\section{Underfitting and overfitting}

\pb{explain title}

\subsection{Underfitting: the universal model}

@@

\pb{What happens if we just look at the witnesses to Komogorov complexity.}
\pb{
stelling: Bounded door universeel model \\
literatuur: Gell-man, Vitanyi noemen univ model.
oplossing: beperk de model klasse
}

\subsection{Overfitting: the singletons}

\pb{
stelling: singletons kunnen domineren in structure function.
lit: Kolmogorov, Vitanyi, Antunes, Gell-Man
oplossing: Paul's meaningful information?
}

\section{The problem of depth}

@@

\begin{theorem}
There are infinitely many $x$ with 
\begin{enumerate}
  \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
  \item $S^\T_{\tau^\circ}(x) \geq K(x)/6$. \label{eq:hondevoer}
\end{enumerate}
\end{theorem}
\begin{proof}
We continue from the proof of Theorem~\ref{theorem:structure-function-is-not-bounded}. Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$. Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{theorem:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$. Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
\end{proof}
Note that we favoured simplicity of the proof over the strength of the bounds: the multiplicative constant $1/6$ can be improved with a more detailed treatment.

\pb{
stelling: voor alle niet superdiepe strings is Paul's sophistication bounded.
}

\section{A world without sophistication}

@@

\pb{Procedural view: we cannot between dominating models. This may not be the aim of algorithmic statistics, but it suggests that it's kind of unreasonable to ask for model selection without repeated samples, or assumptions on the model class.} 
\pb{The problem with making a fixed choice.}
\pb{The existence of balanced strings.}
\pb{Total functions. They ruin efficiency and push sophistication into a land of unprovable claims.}
\pb{reference Antunes instability}

@@

\section{=======Stuff to move up or delete=======}

\section{Old structure}

\subsection{Koppel and Atlan}

We start with sophistication as originally introduced by Koppel and Atlan \cite{koppelSoph1988,koppel1991almost}. Their sophistication is built on a two-argument universal Turing Machine $U(i, y) = \phi_p(y)$. They consider a pair $(i, y)$ a representation of a string $x$ if (a) $\phi_i$ is total, and (b) $\phi_i(y)$ has $x$ as a prefix. The length of a representation is measured as $|i| + |y|$. Note that this ignores the cost of delimiting the model $i$ from the input $y$, ie. this is not the length of a single unambiguous code word for $x$. Candidates are those models with representations $c$ bits from the minimum length, where $c$ is a pre-set `significance' parameter.

\footnotemark

Koppel and Atlan actually provide a machine-independence result, but only in the setting of infinite strings.

This construction shows one of the first pitfalls of sophistication: using an inefficient representation for the model:

Subsequent publications suggest that all ensemeble\ldots.

\begin{theorem}[Reduction of Effective Complexity to Structure
  Function]
  There is a constant $c$ such that, for every representation of $x$
  consisting of an ensemble $E$ with $x\in E$ with entropy $H$,
  yielding a total information of $K(E)+H$, there is a finite set $S$
  with $x\in S$ such that $K(S)\le K(E)+c$ and $\log|S|\le H$.
\end{theorem}
\begin{proof}
  First note that although the ensemble $E$ may itself be infinite, it
  is ``typical'', and therefore a valid candidate representation, for
  only a finite number of its elements, namely the set $S=\{y\in E:-\log
  P(y)\le H\} = \{y:P(y)\ge 2^{-H}\}$. Since the sum of probabilities must
  converge, this set cannot be infinite: since $1\ge\sum_{y\in S} P(y)\ge
  \sum_{y\in S}2^{-H} = |S|2^{-H}$ so $|S|\le 2^H$.
\end{proof}
